---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
  [![R-CMD-check](https://github.com/tylermorganwall/tree3d/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tylermorganwall/tree3d/actions/workflows/R-CMD-check.yaml)
  <!-- badges: end -->

# tree3d: 3D Tree Models

`tree3d` is an R package that provides highly customizable 3D tree models for data visualization and 3D scene creation. `tree3d` allows you to generate tree models as OBJ files, which can then easily be integrated into your data visualization, animation, or 3D rendering workflows.

## Features

- **Variety of tree models**: Includes both planar and solid tree models. 

- **Multiple crown types**: Provides a wide range of crown types, including: columnar, oval, palm, pyramidal, rounded, spreading, vase, and weeping. Each type offers a unique visual aesthetic to represent different types of trees.

- **Customizable dimensions and color**: Allows you to adjust the diameter, height, and color of both the tree's crown and trunk individually. This flexibility allows you to customize the tree model for your specific use case.

## Installation

`tree3d` can currently be installed from Github.

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("tylermorganwall/tree3d")
```

## Examples

Let's load the sandbox data included in the Github repository. We will then unzip the file, and then load the shapefiles into R using {sf}.

```{r}

library(tree3d)
library(sf)
library(terra)
library(rayshader)

#Download the sample data
temp_download = tempfile(fileext = ".zip")
download.file("https://github.com/tylermorganwall/tree3d/raw/main/inst/extdata/tree3d_sandbox_data.zip",temp_download)

#Unzip and get directory
unzip(temp_download, exdir = tempdir())
zipdir = paste0(tempdir(), .Platform$file.sep, 
                "sandbox_data", .Platform$file.sep)

#Load all files
poly_boundary = sf::st_read(paste0(zipdir,"Area_Boundary_subset.shp"), quiet = TRUE)
baseline_tree = sf::st_read(paste0(zipdir,"Baseline_Trees_Z_clip.shp"), quiet = TRUE)
building_baseline = sf::st_read(paste0(zipdir,"Buildings_Baseline_2020_Z_clip.shp"), quiet = TRUE)
parcels = sf::st_read(paste0(zipdir,"Parcels_2020_clip.shp"), quiet = TRUE)

dem = terra::rast(paste0(zipdir,"DEM_2013_subset_zero.tif"))

#Inspect the tree data
head(baseline_tree)
```

Now that our data is loaded, let's generate a 3D map with rayshader. We will load in our 

```{r}
#Use rayshader to convert the DEM to a matrix and generate a 3D mesh
dem_mat = raster_to_matrix(dem)
dem_mat |>
  constant_shade("#6dbd83") |>
  add_overlay(generate_polygon_overlay(poly_boundary, extent = ext(dem),
                                       heightmap = dem_mat,
                                       palette = "#bcebc9"),
              rescale_original = TRUE) |>
  add_overlay(generate_polygon_overlay(parcels, extent = ext(dem),
                                       heightmap = dem_mat,
                                       palette = "#afe0bd"),
              rescale_original = TRUE) |>
  plot_3d(dem_mat, background = "lightblue", windowsize = 800)

#Render the buildings as extruded polygons
render_polygons(building_baseline, extent = ext(dem),
                data_column_top = "height_ag",
                clear_previous = TRUE,
                color = "white",
                heightmap = dem_mat)
render_camera(theta=44, phi=36,  zoom=0.7, fov=50, shift_vertical = -30)
render_snapshot()
```

Now that we have our 3D map generated, let's add our custom trees using rayshader's `render_trees()` function. {tree3d} provides `get_crown_file()` and `get_trunk_file()`, which allows access to the separate 3D models included in the package. Passing in each model separately gives you full control over the size, shape, and colors of the trees.

```{r}

#Render the trees, using data from `baseline_tree` to draw each tree individually
for(i in seq_len(nrow(baseline_tree))) {
  single_tree_info = baseline_tree[i,]
  latlongheight = as.matrix(st_geometry(single_tree_info)[[1]])
  render_tree(long = latlongheight[1],lat = latlongheight[2],
              extent = ext(dem), heightmap = dem_mat,
              crown_height = single_tree_info$Diameter-1,
              crown_width_ratio = 1,
              trunk_height = 1,
              custom_obj_crown = get_crown_file(tree = single_tree_info$CrownType),
              custom_obj_trunk = get_trunk_file())
}
render_camera(40, 23,  0.25, 90)
render_snapshot()

```

## Acknowledgements


