---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tree3d: 3D Tree Models

`tree3d` is an R package that provides highly customizable 3D tree models for data visualization and 3D scene creation. `tree3d` allows you to generate tree models as OBJ files, which can then easily be integrated into your data visualization, animation, or 3D rendering workflows.

## Features

- **Variety of tree models**: Includes both planar and solid tree models. 

- **Multiple crown types**: Provides a wide range of crown types, including: columnar, oval, palm, pyramidal, rounded, spreading, vase, and weeping. Each type offers a unique visual aesthetic to represent different types of trees.

- **Customizable dimensions and color**: Allows you to adjust the diameter, height, and color of both the tree's crown and trunk individually. This flexibility allows you to customize the tree model for your specific use case.

## Installation

`tree3d` can currently be installed from Github.

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("tylermorganwall/tree3d")
```

## Examples

## Acknowledgements


